---
title: "Attrition weighting and influence diagnostics in WFC 0.15"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Attrition weighting and influence diagnostics in WFC 0.15}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
```

WFC 0.15 adds two production diagnostics for recurring surveys and panels:
attrition weighting for wave-to-wave retention, and high-influence unit
diagnostics for reviewing extreme calibrated weights. It also adds Fay's BRR to
the existing replicate-weight variance workflow.

## Weight retained panel units

`wf_attrition()` models whether each prior-wave unit stayed in the next wave.
It returns inverse-retention weights for retained units only. The result is a
`wf_weights` subclass, so it can be composed with later calibration stages.

```{r}
library(WFC)

panel <- data.frame(
  id = paste0("p", 1:12),
  region = rep(c("A", "B"), each = 6),
  retained = c(TRUE, TRUE, TRUE, FALSE, FALSE, FALSE,
               TRUE, TRUE, FALSE, TRUE, FALSE, FALSE),
  age = c(25, 35, 45, 55, 65, 75, 24, 34, 44, 54, 64, 74),
  sex = rep(c("female", "male"), 6)
)

attrition <- wf_attrition(
  panel,
  retained = "retained",
  formula = ~ age + sex,
  id = "id",
  by = "region",
  trim = 3
)

attrition
head(attrition$balance)
```

The `$balance` table compares retained units with the full prior wave before
and after inverse-retention weighting. `wf_report()` carries those diagnostics
as attrition-specific sections.

```{r}
names(wf_report(attrition)$sections)
```

## Review high-influence units

`wf_influence()` ranks units by their weight ratio to the group mean and their
share of squared-weight mass. For the top units, it recomputes the group design
effect after dropping the unit. When sample and target data are provided, it
also reports the largest target-margin share represented by that respondent.

```{r}
data("wfc_example", package = "WFC")

target <- wf_target_population(
  wfc_example$population,
  key_map = c(gender = "gender", age = "age"),
  count = "count",
  dims = wfc_example$dims,
  by = "province"
)
weights <- wf_rake(wfc_example$sample, target, id = "id")

influence <- wf_influence(
  weights,
  target = target,
  sample = wfc_example$sample,
  id = "id",
  top = 5
)

influence
```

Use the table before trimming or category-collapse decisions. It identifies
which individual weights are driving design-effect pressure and which target
margins they dominate.

## Use Fay's BRR

For BRR designs, `rho` switches from standard half-sample BRR to Fay's BRR.
`rho = 0` is the default and reproduces the original WFC behavior.

```{r}
design_data <- data.frame(
  id = paste0("u", 1:8),
  stratum = c("A", "A", "A", "A", "B", "B", "B", "B"),
  psu = c("a1", "a1", "a2", "a2", "b1", "b1", "b2", "b2"),
  y = c(1, 0, 1, 1, 0, 0, 1, 0)
)

refit <- function(data, weights) {
  structure(
    list(
      data = data.frame(
        id = data$id,
        group = "all",
        weight = weights,
        feature = 1 / weights
      )
    ),
    class = "wf_weights"
  )
}

fay <- wf_replicates(
  design_data,
  refit,
  method = "brr",
  strata = "stratum",
  clusters = "psu",
  id = "id",
  rho = 0.3
)

fay
wf_variance(fay, function(w, data) sum(w * data$y) / sum(w), design_data)
```

The stored scale uses the Fay adjustment, so downstream `wf_variance()` and
`as_svrepdesign()` consume the same replicate metadata.
