---
title: "Bringing your own population data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bringing your own population data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Population files rarely use the same column names as a survey. WFC handles this
at the boundary through an explicit key map; the calibration engine only sees a
validated canonical target.

## Map an arbitrary schema

```{r}
library(WFC)
data("wfc_example", package = "WFC")

sample <- wfc_example$sample
sample$region_code <- sample$province
population <- wfc_example$population
names(population) <- c("region_code", "sex_key", "age_band", "n_people")

dims <- wf_dims(
  gender = c("female", "male"),
  age = c("young", "old")
)

target <- wf_target_population(
  population,
  key_map = c(gender = "sex_key", age = "age_band"),
  count = "n_people",
  dims = dims,
  by = "region_code"
)

target$dims
vapply(target$groups, function(group) group$total, numeric(1))
```

`key_map` names are WFC dimension names and its values are source column names.
Category keys are compared as character values. Counts must be finite,
non-missing, and non-negative.

## Choose the target scale

Population totals are the default. For a sample-scale calibration, WFC keeps
the population shares but rescales each group total to its sample size.

```{r}
sample_scale <- wf_target_population(
  population,
  key_map = c(gender = "sex_key", age = "age_band"),
  count = "n_people",
  dims = dims,
  by = "region_code",
  scale = "sample",
  sample = sample
)

vapply(sample_scale$groups, function(group) group$total, numeric(1))
```

Use `scale = "custom"` with named `totals` when another reporting total is the
authority.

## Retain joint cells for post-stratification

Raking needs margins. Post-stratification additionally needs the joint
population cells, requested explicitly with `keep_joint = TRUE`.

```{r}
joint_target <- wf_target_population(
  population,
  key_map = c(gender = "sex_key", age = "age_band"),
  count = "n_people",
  dims = dims,
  by = "region_code",
  keep_joint = TRUE
)

names(joint_target$joint)
head(joint_target$joint[[1]])
```

Do not reshape population data by hand after constructing a target. Rebuild the
target from the authoritative source so its totals, margins, joint cells, and
metadata remain consistent.
