---
title: "Dirty data, missing values, and infeasibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dirty data, missing values, and infeasibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

WFC separates fixable data quality warnings from mathematical infeasibility.
The precheck identifies the group, dimension, and category before any
calibration iteration begins.

```{r}
library(WFC)
data("wfc_example", package = "WFC")

sample <- wfc_example$sample
population <- wfc_example$population
```

## Unknown categories are blocking schema problems

```{r}
target <- wf_target_population(
  population,
  key_map = c(gender = "gender", age = "age"),
  count = "count",
  dims = wfc_example$dims,
  by = "province"
)

unknown <- sample
unknown$gender[1] <- "not-in-target"
unknown_check <- wf_precheck(unknown, target, id = "id")
unknown_check$issues[
  unknown_check$issues$check == "cat_unknown_in_sample",
  c("group", "dim", "category", "severity", "detail")
]
```

Recode the source value or rebuild the target. WFC does not silently drop the
row or invent a target share.

## Missing values require an explicit policy

```{r}
with_na <- sample
with_na$age[1] <- NA

fractional_check <- wf_precheck(with_na, target, id = "id", na = "fractional")
drop_check <- wf_precheck(with_na, target, id = "id", na = "drop")
error_check <- wf_precheck(with_na, target, id = "id", na = "error")

c(
  fractional = fractional_check$ok,
  drop = drop_check$ok,
  error = error_check$ok
)
```

`"fractional"` allocates a missing dimension according to target shares,
`"drop"` removes incomplete rows with a warning, and `"error"` blocks the run.

## Resolve zero support with a declared collapse

An infeasible category has positive target population but no sample support.
Iteration cannot solve it. Declare acceptable merges in `wf_dims()`, let the
precheck generate a reviewable plan, apply it to both sample and target, and
run the precheck again.

```{r}
dims_with_ladder <- wf_dims(
  gender = c("female", "male"),
  age = c("young", "old"),
  .collapse = list(
    age = list(broad = c(young = "all", old = "all"))
  )
)

ladder_target <- wf_target_population(
  population,
  key_map = c(gender = "gender", age = "age"),
  count = "count",
  dims = dims_with_ladder,
  by = "province"
)

sparse <- sample[!(sample$province == "North" & sample$age == "old"), ]
blocked <- wf_precheck(sparse, ladder_target, id = "id")
plan <- wf_suggest_collapse(blocked, dims_with_ladder)
plan

resolved <- wf_apply_collapse(sparse, ladder_target, plan)
recheck <- wf_precheck(resolved$sample, resolved$target, id = "id")
recheck
```

Only a declared ladder can produce an automatic suggestion. If no declared
mapping resolves the issue, WFC leaves it in `plan$unresolved` for human review.
