---
title: "Survey and broom interoperability in WFC 0.12"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Survey and broom interoperability in WFC 0.12}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
has_survey <- requireNamespace("survey", quietly = TRUE)
has_generics <- requireNamespace("generics", quietly = TRUE)
```

WFC 0.12 keeps calibration and downstream analysis separate. WFC constructs
and checks weights; the survey ecosystem estimates means, regressions, totals,
and tables. Bridges align by exact unit ID, so data may be reordered but cannot
silently lose or gain rows.

## Prepare calibrated weights

```{r}
library(WFC)
data("wfc_example", package = "WFC")

sample <- wfc_example$sample
sample$outcome <- as.numeric(sample$age == "young")
target <- wf_target_population(
  wfc_example$population,
  key_map = c(gender = "gender", age = "age"),
  count = "count",
  dims = wfc_example$dims,
  by = "province"
)
weights <- wf_rake(sample, target, id = "id")
```

## Use standard survey estimators

`as_svydesign()` adds `.wf_weight` to a copy of the analysis data and returns a
standard survey design. Cluster, strata, finite-population, and nesting controls
can be supplied when the source design requires them.

```{r survey-design, eval=has_survey}
survey_design <- as_svydesign(weights, sample, id = "id")
survey::svymean(~outcome, survey_design)
```

The result can enter srvyr without another weight join because srvyr accepts
`survey.design2` and `svyrep.design` objects:

```{r srvyr, eval=FALSE}
srvyr_design <- srvyr::as_survey(survey_design)
```

The `survey`, `srvyr`, and their data-manipulation dependencies remain optional;
none is imported by WFC.

## Preserve replicate uncertainty

`as_svrepdesign()` maps WFC bootstrap, unstratified or stratified jackknife, and
BRR metadata to survey's corresponding replication types. It fixes
full-estimate-centered MSE semantics so the survey and WFC variance formulas
agree.

```{r replicate-design, eval=has_survey}
refit <- function(data, base_weights) {
  data$.base_weight <- base_weights
  wf_rake(
    data,
    target,
    id = "id",
    init_weight = ".base_weight",
    precheck = FALSE
  )
}

replicate_weights <- wf_replicates(
  sample,
  refit,
  method = "jackknife",
  id = "id"
)
replicate_design <- as_svrepdesign(
  replicate_weights,
  sample,
  id = "id"
)

wfc_variance <- wf_variance(
  replicate_weights,
  function(w, data) sum(w * data$outcome) / sum(w),
  sample
)
survey_mean <- survey::svymean(~outcome, replicate_design)

c(
  wfc_se = wfc_variance$table$se,
  survey_se = as.numeric(survey::SE(survey_mean))
)
```

## Project results with generics

WFC registers broom-style methods only when the suggested `generics` namespace
is loaded. Every method returns a base data frame with stable English column
names.

```{r tidiers, eval=has_generics}
diagnostics <- wf_diagnose(weights, target)

head(generics::tidy(weights))
generics::glance(diagnostics)

augmented <- generics::augment(weights, data = sample, id = "id")
head(augmented[c("id", ".weight", ".feature")])
```

`tidy()` also projects `wf_diagnostics`, `wf_blend_result`, and
`wf_variance_result`; `glance()` returns a stable one-row summary for each of
those result classes. `augment()` is intentionally limited to `wf_weights`,
where an exact ID join has a clear unit-level meaning.

## Treat ID errors as data-contract failures

Both bridges and `augment()` require unique, non-missing IDs and identical ID
sets between the WFC result and analysis data. Missing, extra, or duplicated
units raise `wf_error_schema`. Existing reserved output columns also raise an
error rather than being overwritten. These checks preserve the same discipline
as calibration precheck at the ecosystem boundary.
