---
title: "Guided workflows and localized output in WFC 0.11"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guided workflows and localized output in WFC 0.11}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
```

`wf_autoweigh()` is a guided path over WFC's existing expert workflow. It uses
the same target constructor, precheck, declared collapse plans, calibration
engines, diagnostics, and reports. Its additional responsibility is to make
each automated choice explicit and auditable.

## Run the guided raking path

The bundled data are simulated. Passing raw population data lets the guided
workflow construct the canonical target before enforcing precheck.

```{r}
library(WFC)
data("wfc_example", package = "WFC")

guided <- wf_autoweigh(
  sample = wfc_example$sample,
  population = wfc_example$population,
  dims = wfc_example$dims,
  key_map = c(gender = "gender", age = "age"),
  count = "count",
  by = "province",
  id = "id",
  caps = c(2, 4, 8),
  interactive = FALSE
)

guided$weights
guided$diagnostics
```

In non-interactive mode, only pre-declared and reproducible remediations may be
applied. A blocking issue without an applicable map in `wf_dims()` raises a
classed feasibility error before weights are produced.

## Audit the decision ledger

The ledger is an ordered data frame. `action` and `detail_key` are stable
English programmatic keys; `detail` is localized human-facing narration.
`artifacts` is aligned one-for-one with ledger rows and retains objects such as
the target, precheck, collapse plan, trim frontier, and final report.

```{r}
guided$ledger[c(
  "step", "action", "detail_key", "artifact_class"
)]

length(guided$artifacts) == nrow(guided$ledger)
```

Keep the result with any published estimate. It contains the final sample and
target after declared remediation, the selected method, normalized language,
weights, diagnostics, report, ledger, and structured artifacts.

## Understand method selection

`method = "auto"` chooses raking unless both a reviewed
`wf_collapse_ladder` and a positive `min_cell` are supplied. With both inputs it
chooses post-stratification and retains joint population cells when building a
raw target. Automatic mode never chooses bounded logit calibration because
weight bounds require analyst judgment.

Use explicit `method = "logit"` with reviewed bounds:

```{r}
logit_result <- wf_autoweigh(
  sample = wfc_example$sample,
  population = guided$target,
  dims = wfc_example$dims,
  id = "id",
  method = "logit",
  bounds = c(0.3, 3),
  trim = NULL,
  interactive = FALSE
)

logit_result$method
```

## Localize human-facing output

English and Simplified Chinese are supported. Resolution follows an explicit
`lang` argument, then `options(wfc.lang)`, then the session locale, with English
as the fallback. Unsupported explicit values raise `wf_error_input`.

```{r}
localized <- wf_report(
  guided$weights,
  guided$target,
  audience = "manager",
  lang = "zh_CN"
)

localized$language
identical(names(localized$table), names(guided$report$table))
```

Localization does not translate error messages, R documentation, object or
column names, condition classes, user data, ledger actions, or detail keys.
This keeps automated consumers stable across languages.

## Preserve review boundaries

The guided path accelerates routine execution but does not replace statistical
review. Analysts still own dimension definitions, allowed collapse maps,
post-stratification ladders, minimum cell sizes, logit bounds, trim criteria,
and publication decisions. Use `interactive = TRUE` when each declared collapse
or finite trim recommendation should require confirmation before execution.
