---
title: "Soft calibration and entropy balancing in WFC 0.14"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Soft calibration and entropy balancing in WFC 0.14}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
```

WFC 0.14 adds two calibration methods behind the existing `wf_calibrate()`
interface. `method = "soft"` is for declared, auditable relaxation when exact
margin matching is not defensible or not feasible. `method = "ebal"` is for
entropy balancing: exact constraints with minimum divergence from the base
weights.

## Relax margins deliberately

Soft calibration keeps group totals exact and allows margin gaps only within
declared tolerances. The tolerance is a fraction of the group total and may be a
single value or a named vector by dimension.

```{r}
library(WFC)

dims <- wf_dims(segment = c("covered", "missing"))
target <- wf_target_manual(
  data.frame(
    dimension = c("segment", "segment"),
    category = c("covered", "missing"),
    value = c(90, 10)
  ),
  dims
)
sample <- data.frame(
  id = paste0("s", 1:4),
  segment = "covered"
)

soft <- wf_calibrate(
  sample,
  target,
  method = "soft",
  id = "id",
  tolerance = 0.11
)

soft
soft$relaxation[c(
  "group", "dim", "category", "target", "achieved",
  "relative_gap", "within_tolerance"
)]
```

If the same zero-support category is outside tolerance, WFC refuses to produce
weights. That keeps soft calibration explicit rather than silently hiding an
unusable target.

```{r eval=FALSE}
wf_calibrate(
  sample,
  target,
  method = "soft",
  id = "id",
  tolerance = 0.05
)
```

`wf_report()` includes the same relaxation table, so the published report can
show which margins were relaxed and by how much.

## Match moments exactly with entropy balancing

Entropy balancing uses exponential weights to satisfy the standard categorical
target and optional continuous moment targets. Moment values are target means,
not totals; WFC multiplies them by the group total internally.

```{r}
dims2 <- wf_dims(gender = c("female", "male"))
target2 <- wf_target_manual(
  data.frame(
    dimension = c("gender", "gender"),
    category = c("female", "male"),
    value = c(2, 2)
  ),
  dims2
)
sample2 <- data.frame(
  id = paste0("e", 1:4),
  gender = c("female", "female", "male", "male"),
  x = c(0, 1, 0, 1)
)

ebal <- wf_calibrate(
  sample2,
  target2,
  method = "ebal",
  id = "id",
  moments = c(x = 0.75)
)

ebal
ebal$moments
```

The categorical margin and the continuous target mean are both exact within the
solver tolerance:

```{r}
sum(ebal$data$weight[sample2$gender == "male"])
sum(ebal$data$weight * sample2$x) / sum(ebal$data$weight)
```

Use entropy balancing when exact moment matching is the requirement. Use soft
calibration when the target itself requires a declared tolerance and the
relaxation audit table is part of the deliverable.
