---
title: "The precheck-execute-diagnose workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The precheck-execute-diagnose workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

WFC treats weighting as a reviewable workflow, not a single optimizer call. A
normal analysis declares the dimensions, builds one canonical target, checks
feasibility, runs a calibration engine, and inspects the result.

## Declare dimensions and build a target

The bundled data are fully simulated and contain no private survey records.

```{r}
library(WFC)
data("wfc_example", package = "WFC")

sample <- wfc_example$sample
population <- wfc_example$population
dims <- wfc_example$dims

target <- wf_target_population(
  population,
  key_map = c(gender = "gender", age = "age"),
  count = "count",
  dims = dims,
  by = "province"
)
```

Every external source becomes the same `wf_target` representation. Downstream
engines therefore do not depend on the original population-table schema.

## Precheck before execution

```{r}
check <- wf_precheck(sample, target, id = "id")
check
```

Errors block execution. Warnings and notes remain visible and are carried into
the audit trail; there is no `force = TRUE` bypass for infeasible calibration.

## Execute and diagnose

```{r}
weights <- wf_rake(sample, target, id = "id")
weights

diagnostics <- wf_diagnose(weights, target = target)
diagnostics
diagnostics$table
```

The returned `wf_weights` object contains unit weights, convergence logs,
achieved margins, and provenance. Keep that object with the published estimate:
it records the settings and package version needed to reconstruct the run.

## Use the unified dispatcher

The same target can be passed through the common interface. This returns the
same `wf_weights` contract as the direct engine call.

```{r}
via_dispatcher <- wf_calibrate(
  sample,
  target,
  method = "raking",
  id = "id"
)

all.equal(
  weights$data$weight,
  via_dispatcher$data$weight,
  tolerance = 1e-8
)
```
