---
title: "Production infrastructure in WFC 0.13"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Production infrastructure in WFC 0.13}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
```

WFC 0.13 adds a production layer over the existing engines. The layer does not
change raking, post-stratification, propensity correction, or diagnostics. It
records how a recurring weighting round should be run, executes that declared
round, checks whether new weights drifted from a reference release, and writes
machine-readable audit records.

## Declare a pipeline

The pipeline specification is a plain list with a stable hash. Runtime data are
passed to `wf_run()`, so the spec can be versioned separately from each survey
wave.

```{r}
library(WFC)
data("wfc_example", package = "WFC")

spec <- wf_pipeline(
  target = list(
    mode = "population",
    key_map = c(gender = "gender", age = "age"),
    count = "count",
    by = "province"
  ),
  stages = list(
    calibrate = list(method = "raking", id = "id", tol = 1e-8)
  ),
  validate = list(max_deff = 6, max_margin_dev = 1e-4)
)

spec
```

The same declaration can use `mode = "reference"`, `mode = "manual"`, or a
ready `wf_target` object when the target is already constructed by an upstream
review step.

## Run the declared round

`wf_run()` builds the target, executes optional pre-calibration stages, runs the
calibration stage, and attaches pipeline provenance to the returned
`wf_weights`.

```{r}
weights <- wf_run(
  spec,
  sample = wfc_example$sample,
  dims = wfc_example$dims,
  population = wfc_example$population
)

weights$provenance$pipeline[c("hash", "target_mode")]
weights$pipeline_validation$ok
```

Numeric `base_weight` values are accepted so the same pipeline can be used as a
`wf_replicates()` refit body:

```{r eval=FALSE}
refit <- function(data, base_weights) {
  wf_run(
    spec,
    sample = data,
    dims = wfc_example$dims,
    population = wfc_example$population,
    base_weight = base_weights
  )
}
```

An optional `propensity` stage may be declared before `calibrate`; its
pseudo-weights are multiplied into the calibration initial weights.

## Check drift against a reference release

`wf_validate()` compares new weights with a reference release. It checks group
coverage, design effect, effective sample size, total weight, and matched-unit
weight-ratio drift. When a target is supplied, margin residual drift is checked
as well.

```{r}
reference <- weights
new <- weights
new$data$weight[1] <- new$data$weight[1] * 3

drift <- wf_validate(
  new,
  reference,
  max_deff_delta = 0.01,
  max_ratio_p99 = 1.1,
  on_issue = "none"
)

drift
drift$issues[c("check", "group", "severity")]
```

Use `on_issue = "warn"` in automated monitoring to emit a classed
`wf_warning_quality`, or `on_issue = "error"` when drift should stop the
pipeline.

## Export an audit record

`wf_audit_export()` writes dependency-free JSON. The record contains result
provenance, pipeline metadata, summary weight statistics, optional guided
workflow ledgers, hashes of named inputs, and user-supplied metadata.

```{r}
audit_file <- tempfile(fileext = ".json")
wf_audit_export(
  weights,
  audit_file,
  inputs = list(
    sample = wfc_example$sample,
    population = wfc_example$population
  ),
  extra = list(round = "example")
)

file.exists(audit_file)
```

Keep the audit record with published estimates. It is the machine-readable
counterpart to the human-facing `wf_report()` output and the guided workflow
decision ledger.
