---
title: "Usability foundations in WFC 0.10"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Usability foundations in WFC 0.10}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

WFC 0.10 adds recommendation, reporting, and visualization tools around the
existing calibration engines. They preserve the package's review discipline:
recommendations are artifacts for approval, not automatic mutations.

## Prepare one calibrated result

```{r}
library(WFC)
data("wfc_example", package = "WFC")

sample <- wfc_example$sample
dims <- wfc_example$dims
target <- wf_target_population(
  wfc_example$population,
  key_map = c(gender = "gender", age = "age"),
  count = "count",
  dims = dims,
  by = "province"
)
weights <- wf_rake(sample, target, id = "id")
diagnostics <- wf_diagnose(weights, target)
```

## Recommend a trim cap

`wf_auto_trim()` re-runs raking for each candidate and an untrimmed baseline.
It reports design effect, margin distortion, warnings, and candidate failures.
The returned cap is not applied automatically.

```{r}
trim_advice <- wf_auto_trim(
  sample,
  target,
  id = "id",
  caps = c(2, 4, 8),
  max_deff = 6,
  max_residual = 0.02
)
trim_advice
```

```{r trim-plot, fig.width=7, fig.height=3.5}
plot(trim_advice)
```

## Draft a collapse ladder

The draft uses explicit level order from `wf_dims()` and worst-group support.
A deliberately high threshold below forces a small demonstration draft. Review
the mappings before using `draft$ladder` in post-stratification.

```{r}
draft <- wf_suggest_ladder(sample, target, dims, min_cell = 25)
draft
draft$ladder
```

## Build a manager or analyst dossier

Both audiences read one structured object. The manager projection adds traffic
lights and publication actions; the analyst projection retains the full
diagnostic table and method-specific sections.

```{r}
manager_report <- wf_report(weights, target, audience = "manager")
manager_report

analyst_report <- wf_report(weights, target, audience = "analyst")
as.data.frame(analyst_report)
```

Markdown and standalone escaped HTML use the same payload:

```{r}
markdown <- wf_report(weights, target, output = "markdown")
cat(substr(markdown, 1, 240), "...\n")
```

Version 0.11 activates `lang = "en"` and `lang = "zh_CN"` for reports and plot
labels. Localization changes only human-facing rendering; stable object fields,
condition classes, and diagnostic values remain unchanged.

## Inspect the result visually

```{r result-plots, fig.width=7, fig.height=3.5}
plot(weights, max_groups = 2)
plot(diagnostics)
```
